Nonplanarity and Street Network Representation in Urban Form Studies

### 2018-01-24 manuscript.tex
add explanatory figure and edits throughout

\runninghead{Author Redacted}
\title{The Importance of Being Nonplanar: Street Network Representation in Urban Form Studies}
\author{Author Redacted \affilnum{1}}
\affiliation{\affilnum{1}Affiliation redacted}
\corrauth{Author Redacted, Address Redacted}
\email{Email Redacted}

@@

@@

In urban planning and design research, street networks are routinely used to calculate accessibility between origins and destinations or to compute indicators of the urban form, such as block sizes or intersection density and connectivity. Mathematical models of street networks, called graphs, have grown ubiquitous in the urban studies literature in recent years as they have been used to model household travel patterns, access and equity, pedestrian volume, urban design patterns and spatial morphology, and location centrality and polycentricity \citep{marshall_street_2010,porta_alterations_2014,marshall_community_2014,hajrasouliha_impact_2015,parthasarathi_street_2015,knight_metrics_2015,gil_street_2016,zhong_revealing_2017}.

@@

Consider two theoretical examples: a medieval European city center and a modern American city center. The former's circulation network comprises a set of pedestrian paths and drivable streets without any grade separation, overpasses, or underpasses. In two-dimensions, its edges (i.e., streets and paths) cross each other only at nodes (i.e., intersections), so it is by definition planar. The latter's circulation network comprises pedestrian paths, drivable surface streets, and a grade-separated freeway with overpasses over some of the surface streets. In two-dimensions, its edges occasionally cross each other at non-nodes (i.e., overpasses), so it is by definition nonplanar. However, we might call it \emph{approximately} planar because its nonplanar edge crossings are relatively uncommon. Approximate planarity constrains nonplanar spatial networks such that they do not exhibit certain characteristics found among nonplanar aspatial graphs (e.g., small-world effects or power-law distributed node degrees).

@@

The graph theory literature offers some measures of how \enquote{far off} a nonplanar graph $G$ is from being planar, including its \emph{crossing number} --- the minimum number of edge crossings of any drawing of $G$ --- and its \emph{skewness} --- the minimum number of edges that must be removed from $G$ to produce a planar graph \citep{liebers_planarizing_2001,szekely_successful_2004,chimani_non-planar_2009}. However, these measures are imperfect and hard to compute \citep{chimani_vertex_2012,szekely_successful_2004}. They also fail to correct for the size or density of a spatial network. In his discussion of road networks and approximate spatial planarity, \citet[p.~133]{newman_networks:_2010} argues that due to these drawbacks \enquote{no widely accepted metric for degree of planarity has emerged,} and calls for the development of better indicators.

@@

To acquire these street networks, we use OSMnx to download the data for each city and network type from OpenStreetMap. OpenStreetMap is a collaborative online mapping platform commonly used by researchers because of its good worldwide coverage \citep{haklay_how_2010,jokar_arsanjani_openstreetmap_2015}. OSMnx is a Python-based software tool that allows us to automatically download a street network from OpenStreetMap for any study site in the world, and automatically process it into a length-weighted nonplanar directed graph \citep{boeing_osmnx:_2017}. It differentiates between walkable and drivable routes in the circulation network based on individual elements' metadata that describe how the route may be used. Thus the walkable network may contain surface streets, paths through parks, pedestrian passageways between buildings or under roads, and other walkable paths. The drivable network may contain surface streets, grade-separated freeways, and other drivable paths.

@@

Once we have acquired and prepared our networks, we calculate three measures of planarity. The first is an aspatial test of formal planarity using the algorithm described by \citet{boyer_subgraph_2012}. This assesses if it is possible to rearrange the graph's nodes and edges in space, while preserving its topology, so that edges cross only at nodes. This binary true/false indicator tells us if the graph is formally planar, ignoring its real-world spatial embedding. However, street networks \emph{are} spatially embedded, so we have developed the following two indicators to assess the \enquote{extent} to which they are planar.

The second measure is the spatial planarity ratio (SPR). It represents the ratio of nonplanar intersections $i_n$ (i.e., non-dead-end nodes in the nonplanar, three-dimensional, spatially-embedded graph) to planar intersections $i_p$ (i.e., line crossings in the planar, two-dimensional, spatially-embedded graph):

\begin{equation}
	\label{eq:spr}
	ELR = \frac{i_n}{i_p}
\end{equation}

This indicates the extent to which planarity overstates intersections and connectivity in each street network. In other words, the SPR tells us what proportion of the line crossings in the planar graph are true intersections in the nonplanar graph. A truly planar network with no bridges or tunnels will thus have an SPR score of 1.0, while lower values indicate the percentage extent to which the network is planar. From the SPR, we can further calculate by what percentage the planar graph overstates intersection counts as:

\begin{equation}
	\label{eq:spr_overstates}
	1 - \frac{1}{SPR}
\end{equation}

The third measure is the edge length ratio (ELR). It represents the ratio of the mean edge length in the planar graph $l_p$ to the mean edge length in the nonplanar graph $l_n$:

\begin{equation}
	\label{eq:elr}
	ELR = \frac{l_p}{l_n}
\end{equation}

This indicates the extent to which planarity understates edge lengths in each street network by fragmenting street segments at overpasses or underpasses. A truly planar street network with no bridges or tunnels will thus have an ELR score of 1.0, while lower values indicate the percentage extent to which the network is planar. From the ELR, we can further calculate by what percentage the planar graph understates the average edge length as:

\begin{equation}
	\label{eq:elr_understates}
	1 - ELR
\end{equation}

@@

\begin{table*}[htbp]
	\centering
	\caption{Planarity measures for the central street networks of 50 cities worldwide (Planar = whether street network passed the formal test of planarity; SPR = spatial planarity ratio; ELR = edge length ratio).}
	\label{tab:world_cities}
	\input{table_world_cities.tex}
\end{table*}

\begin{figure*}[htbp]
	\center
	\includegraphics[width=\textwidth]{world_map_phi_bw.png}
	\caption{Map of world cities from Table \ref{tab:world_cities} grouped by SPR score terciles (lower scores mean less planar).}
	\label{fig:world_map_bw}
\end{figure*}

\begin{figure}[htbp]
	\includegraphics[width=0.48\textwidth]{regression_phi_split.png}
	\caption{Log-log plot of ELR vs SPR, by network type, with simple regression lines.}
	\label{fig:regression_split}
\end{figure}

Table \ref{tab:world_cities} shows the SPR, ELR, and results of the formal planarity test for the walkable and the drivable networks for each of the 50 cities. Among the drivable street networks, only 20\% are formally planar. On average, they are 88.5\% planar by the SPR measure and 90\% planar by the ELR measure. The individual SPR values indicate that spatial planarity ranges from a high of 100\% in seven of these cities to a low of 57\% in Moscow. The ELR values indicate that spatial planarity ranges from a high of 100\% in seven of these cities to a low of 63.5\% in Los Angeles.

Among walkable street networks, only 10\% are formally planar. On average, they are 92\% planar by the SPR measure and 91.5\% planar by the ELR measure. The individual SPR values indicate that spatial planarity ranges from a high of 100\% in two cities to a low of 67\% in Shanghai. The ELR values indicate that spatial planarity ranges from a high of 100\% in two cities to a low of 66\% in Shanghai. Fewer walkable than drivable networks are formally planar, but on average these walking networks are slightly more spatially planar than the driving networks.

Not all formally planar street networks are spatially planar. For example, Toronto's drivable network is formally planar but only 93\% and 96\% spatially planar according to its SPR and ELR measures. In total, three drivable networks (Toronto, Jakarta, and Copenhagen) and three walkable networks (Dallas, Delhi, and Bologna) are formally planar but spatially nonplanar to various extents. About a third of the cities studied demonstrate spatial planarity of 97\% or higher according to the SPR, suggesting that they are \enquote{approximately} planar. However, another third of the cities studied are less than 87\% planar. Dallas, Los Angeles, and Moscow have SPR values below 60\%, suggesting planar graphs poorly model these city centers. Moreover, planar graphs overstate the intersection counts in these three networks by 67\%, 72\%, and 74\% respectively (Equation \ref{eq:spr_overstates}).

Mogadishu is the only city studied that demonstrates perfect planarity across all three indicators for both network types. All three Italian cities demonstrate perfect planarity in their centers' drivable networks, but not in their walkable networks. The extent of planarity is not consistent across network types: Dallas's walkable SPR is 61\% greater than its drivable SPR, while Geneva's drivable SPR is 19\% greater than its walkable SPR. Figure \ref{fig:world_map_bw} shows the distribution of SPR values around the world. While nearly every European city is in the highest tercile, indicating their networks are more planar, most American cities are in the lowest tercile, indicating their networks are more nonplanar.

Figure \ref{fig:regression_split} shows the relationship between the SPR and ELR indicators across all 50 cities for both network types, as a log-log plot. The indicators' log-log relationship is linear, positive, and very strong (drivable $r^2=0.98$ and walkable $r^2=0.99$). This strong correlation tells us that these two indicators unsurprisingly provide redundant statistical information about the extent to which a network is planar. However, each assesses different implications of this bias for measuring the urban form.

@@

Finally, we examine how these measures behave across an entire city. We find that Oakland's city-wide street network is formally nonplanar. The city has an SPR score of 91.8\% and an ELR score of 93.6\%. This suggests that the planar representation of Oakland's drivable street network overstates the number of intersections --- and thus, the network's connectivity --- by 8.9\% city-wide (Equation \ref{eq:spr_overstates}) and understates the average edge length by 6.4\% city-wide (Equation \ref{eq:elr_understates}).

However, these indicators' values vary across the city. To explore this statistical variation, Table \ref{tab:samples_city} presents summary statistics of these planarity indicators across 100 square-mile samples of Oakland's drivable street network. Our samples' mean SPR and ELR scores are reasonably close to the city-wide values. However, the samples range from spatial planarity lows of 56.9\% (SPR) and 63.7\% (ELR) up to highs of 100\%. 67\% of the samples pass the formal planarity test. However, 63\% of the samples are at least somewhat spatially nonplanar (i.e., with SPR less than 1).

@@

Street networks are frequently spatially nonplanar because they are embedded in three dimensions, not two: they have a z-coordinate (elevation) along with their x- and y-coordinates. But because they are usually \enquote{mostly} planar (average drivable SPR of 88.5\% and average walkable SPR of 92\%), typically with only a few overpasses or underpasses, they could often be described as \emph{approximately planar}. However, claiming that urban street networks universally are \enquote{planar} misrepresents them in several ways:

@@

As George Box famously stated, \enquote{All models are wrong but some are useful.} Even if they are not formally planar, can street networks be simplified to planar graphs and still be usefully well-modeled? Our results suggest that the answer depends on the study site and on the type of analysis. In limited circumstances, where the circulation network and built form exhibit few (or ideally zero) underpasses, overpasses, or grade-separation, then perhaps yes.

@@

We might repurpose this discrepancy between planar and nonplanar representations of a street network to reveal something useful about urban form. The SPR and ELR scores of the major US city centers indicate the greater three-dimensionality of their transportation infrastructure compared to that of the European city centers. These US city centers feature more automobile-oriented roadways and are often bounded by grade-separated freeways. By contrast, these European city centers feature less three-dimensionality as their streets and paths tend to be at-grade. Examining how SPR and ELR change over time in different cities could tell us something about the type and pace of urban development.

It is also interesting that most street networks are formally nonplanar while demonstrating SPR and ELR scores that are only slightly nonplanar. Approximate planarity is common across cities because of costs, technology, and politics. Why? It is expensive to engineer a three-dimensional network with z-axis extents (nearly) as extensive as the x- and y-axes. Furthermore, a city fully consumed by overpasses and tunnels would present degraded livability and walkability, making it politically infeasible.

@@

%\clearpage
\bibliographystyle{apa}
\bibliography{references}

>>>>

\runninghead{Author Redacted}
\title{Nonplanarity and Street Network Representation in Urban Form Studies}
\author{Author Redacted \affilnum{1}}
\affiliation{\affilnum{1}Affiliation redacted}
\corrauth{Author Redacted, Address Redacted}
\email{Email Redacted}

@@

Mathematical models of street networks, called graphs, are widely used throughout the urban studies literature to examine travel behavior, accessibility, urban design patterns, and morphology. These networks are commonly called planar graphs, meaning that they can be represented in two dimensions without any underpasses or overpasses. However, real urban street networks exist in three-dimensional space and frequently include grade separation such as bridges and tunnels. How well do planar graphs model real-world street networks? What does the extent of nonplanarity tell us about a place's infrastructure and urban development? We measure the nonplanarity of drivable and walkable street networks in the centers of 50 cities worldwide, then examine the variation of planarity across a single city. We find that while some street networks are approximately planar, many are poorly-modeled by planar graphs. In particular, planarity can drastically misrepresent routing, connectivity, intersection counts/densities, and block lengths.

@@

In urban planning and design research, street networks are routinely used to calculate accessibility between origins and destinations or to compute indicators of the urban form, such as block sizes or intersection density and connectivity. Mathematical models of street networks, called graphs, have grown ubiquitous in the urban studies literature in recent years as they have been used to model household travel patterns, access and equity, pedestrian volume, urban design patterns and spatial morphology, and location centrality and polycentricity \citep{southworth_street_1995,marshall_street_2010,porta_alterations_2014,marshall_community_2014,hajrasouliha_impact_2015,parthasarathi_street_2015,knight_metrics_2015,gil_street_2016,zhong_revealing_2017}.

@@

\begin{figure}[tbp]
	\center
	\includegraphics[width=0.48\textwidth]{planar_vs_not.png}
	\caption{Formally and spatially planar (left) and nonplanar (right) road networks.}
	\label{fig:planar_vs_not}
\end{figure}

Consider the examples in Figure \ref{fig:planar_vs_not}. The left network comprises a set of paths and streets without any bridges or tunnels. In two-dimensions, its edges intersect each other only at nodes so it is by definition planar. The right network comprises surface streets and a grade-separated freeway interchange with many overpasses and underpasses. In two-dimensions, its edges sometimes cross each other at non-nodes (i.e., overpasses), so it is spatially nonplanar. Furthermore it is impossible to redraw the graph by moving its nodes and edges around in space such that its edges only intersect at nodes. Therefore it is also formally nonplanar. However, at a city- or region-wide spatial scale, these nonplanar edge crossings may be relatively uncommon; we might call such a network \emph{approximately} planar. Approximate planarity constrains nonplanar spatial networks such that they do not exhibit certain characteristics found among nonplanar aspatial graphs, such as small-world effects or power-law distributed node degrees \citep{fischer_spatial_2014}.

@@

The graph theory literature offers some measures of how \enquote{far off} a nonplanar graph $G$ is from being planar, including its \emph{crossing number} --- the minimum number of edge crossings of any drawing of $G$ --- and its \emph{skewness} --- the minimum number of edges that must be removed from $G$ to produce a planar graph \citep{liebers_planarizing_2001,chimani_non-planar_2009}. However, these measures are imperfect and hard to compute \citep{szekely_successful_2004,chimani_vertex_2012}. They also fail to correct for the size, density, or real-world embedding of a spatial network. In his discussion of road networks and approximate spatial planarity, \citet[p.~133]{newman_networks:_2010} argues that due to these drawbacks \enquote{no widely accepted metric for degree of planarity has emerged,} and calls for the development of better indicators.

@@

To acquire these street networks, we use OSMnx to download the data for each city and network type from OpenStreetMap. OpenStreetMap is a collaborative online mapping platform commonly used by researchers because of its good worldwide coverage \citep{haklay_how_2010,jokar_arsanjani_openstreetmap_2015}. OSMnx is a Python-based software tool that allows us to automatically download a street network from OpenStreetMap for any study site in the world, and automatically process it into a length-weighted nonplanar directed graph \citep{boeing_osmnx:_2017}. It differentiates between walkable and drivable routes in the circulation network based on individual elements' metadata that describe how the route may be used. Thus the walkable network may contain surface streets, paths through parks, pedestrian passageways between buildings or under roads, and other walkable paths. The drivable network may contain surface streets, grade-separated freeways, and other drivable routes.

@@

Once we have acquired and prepared our networks, we calculate three measures of planarity. The first is an aspatial test of formal planarity using the algorithm described by \citet{boyer_subgraph_2012}. This assesses if it is possible to rearrange the graph's nodes and edges in space, while preserving its topology, so that edges cross only at nodes. This binary true/false indicator tells us if the graph is formally planar, ignoring its real-world spatial embedding. However, street networks \emph{are} spatially embedded. Accordingly, for this study, we have developed a second and third indicator to assess the \enquote{extent} to which they are planar.

The second measure is the spatial planarity ratio $\phi$. It represents the ratio of nonplanar intersections $i_n$ (i.e., non-dead-end nodes in the nonplanar, three-dimensional, spatially-embedded graph) to planar intersections $i_p$ (i.e., edge crossings in the planar, two-dimensional, spatially-embedded graph):

\begin{equation}
	\label{eq:spr}
	\phi = \frac{i_n}{i_p}
\end{equation}

This indicates the extent to which planarity overstates intersections and connectivity in each street network. In other words, $\phi$ tells us what proportion of the two-dimensional edge crossings in the planar graph are true intersections in the nonplanar graph. A truly planar network with no bridges or tunnels will thus have an $\phi$ score of 1.0, while lower values indicate the percentage extent to which the network is planar. From the $\phi$, we can further calculate by what percentage the planar graph overstates intersection counts as:

\begin{equation}
	\label{eq:spr_overstates}
	1 - \frac{1}{\phi}
\end{equation}

The third measure is the edge length ratio ($\lambda$). It represents the ratio of the mean edge length in the planar graph $l_p$ to the mean edge length in the nonplanar graph $l_n$:

\begin{equation}
	\label{eq:elr}
	\lambda = \frac{l_p}{l_n}
\end{equation}

This indicates the extent to which planarity understates edge lengths in each street network by fragmenting street segments at overpasses or underpasses. A truly planar street network with no bridges or tunnels will thus have an $\lambda$ score of 1.0, while lower values indicate the percentage extent to which the network is planar. From the $\lambda$, we can further calculate by what percentage the planar graph understates the average edge length as:

\begin{equation}
	\label{eq:elr_understates}
	1 - \lambda
\end{equation}

@@

\begin{table*}[htbp]
	\centering
	\caption{Planarity measures for the central street networks of 50 cities worldwide (Planar = whether street network passed the formal test of planarity; $\phi$ = spatial planarity ratio; $\lambda$ = edge length ratio).}
	\label{tab:world_cities}
	\input{table_world_cities.tex}
\end{table*}

\begin{figure*}[htbp]
	\center
	\includegraphics[width=\textwidth]{world_map_phi_bw.png}
	\caption{Map of world cities from Table \ref{tab:world_cities} grouped by $\phi$ terciles (lower values mean less planar).}
	\label{fig:world_map_bw}
\end{figure*}

\begin{figure}[htbp]
	\includegraphics[width=0.48\textwidth]{regression_phi_split.png}
	\caption{Log-log plot of $\lambda$ vs $\phi$, by network type, with simple regression lines.}
	\label{fig:regression_split}
\end{figure}

Table \ref{tab:world_cities} shows $\phi$, $\lambda$, and results of the formal planarity test for the walkable and the drivable networks in these 50 city centers. Among the drivable street networks, only 20\% are formally planar. On average, they are 88.5\% planar by the $\phi$ measure and 90\% planar by the $\lambda$ measure. The individual $\phi$ values indicate that spatial planarity ranges from a high of 100\% in seven of these cities to a low of 57\% in Moscow. The $\lambda$ values indicate that spatial planarity ranges from a high of 100\% in seven of these cities to a low of 63.5\% in Los Angeles. On average across these networks, planar representations overcount intersections by 16\% and underestimate street segment lengths by 10\%.

Among walkable street networks, only 10\% are formally planar. On average, they are 92\% planar by the $\phi$ measure and 91.5\% planar by the $\lambda$ measure. The individual $\phi$ values indicate that spatial planarity ranges from a high of 100\% in two cities to a low of 67\% in Shanghai. The $\lambda$ values indicate that spatial planarity ranges from a high of 100\% in two cities to a low of 66\% in Shanghai. On average across these networks, planar representations overcount intersections by 9\% and underestimate street segment lengths by 8.5\%. Fewer walkable than drivable networks are formally planar, but on average these walking networks are slightly more spatially planar than the driving networks.

Not all formally planar street networks are spatially planar. For example, Toronto's drivable network is formally planar but only 93\% and 96\% spatially planar according to its $\phi$ and $\lambda$ measures. In total, three drivable networks (Toronto, Jakarta, and Copenhagen) and three walkable networks (Dallas, Delhi, and Bologna) are formally planar but spatially nonplanar to various extents. About a third of the cities studied demonstrate spatial planarity of 97\% or higher according to the $\phi$, suggesting that they are \enquote{approximately} planar. However, another third of the cities studied are less than 87\% planar. Dallas, Los Angeles, and Moscow have $\phi$ values below 60\%, suggesting planar graphs poorly model these city centers. Moreover, planar graphs overstate the intersection counts in these three networks by 67\%, 72\%, and 74\% respectively (Equation \ref{eq:spr_overstates}).

Mogadishu is the only city studied that demonstrates perfect planarity across all three indicators for both network types. All three Italian cities demonstrate perfect planarity in their centers' drivable networks, but not in their walkable networks. The extent of planarity is not consistent across network types: Dallas's walkable $\phi$ is 61\% greater than its drivable $\phi$, while Geneva's drivable $\phi$ is 19\% greater than its walkable $\phi$. Figure \ref{fig:world_map_bw} shows the distribution of $\phi$ values around the world. While nearly every European city is in the highest tercile, indicating their networks are more planar, most American cities are in the lowest tercile, indicating their networks are more nonplanar.

Figure \ref{fig:regression_split} shows the relationship between the $\phi$ and $\lambda$ indicators across all 50 cities for both network types, as a log-log plot. The indicators' log-log relationship is linear, positive, and very strong (drivable $r^2=0.98$ and walkable $r^2=0.99$). This strong correlation tells us that these two indicators unsurprisingly provide redundant statistical information about the extent to which a network is planar. However, each assesses different implications of this bias for measuring the urban form.

@@

Finally, we examine how these measures behave across an entire city. We find that Oakland's city-wide street network is formally nonplanar. The city has a $\phi$ score of 91.8\% and an $\lambda$ score of 93.6\%. This suggests that the planar representation of Oakland's drivable street network overstates the number of intersections --- and thus, the network's connectivity --- by 8.9\% city-wide (Equation \ref{eq:spr_overstates}) and understates the average edge length by 6.4\% city-wide (Equation \ref{eq:elr_understates}).

However, these indicators' values vary across the city. To explore this statistical variation, Table \ref{tab:samples_city} presents summary statistics of these planarity indicators across 100 square-mile samples of Oakland's drivable street network. Our samples' mean $\phi$ and $\lambda$ scores are reasonably close to the city-wide values. However, the samples range from spatial planarity lows of 56.9\% ($\phi$) and 63.7\% ($\lambda$) up to highs of 100\%. 67\% of the samples pass the formal planarity test. However, 63\% of the samples are at least somewhat spatially nonplanar (i.e., with $\phi$ less than 1).

@@

Street networks are frequently spatially nonplanar because they are embedded in three dimensions, not two: they have a z-coordinate (elevation) along with their x- and y-coordinates. But because they are usually \enquote{mostly} planar (average drivable $\phi$ of 88.5\% and average walkable $\phi$ of 92\%), typically with only a few overpasses or underpasses, they could often be described as \emph{approximately planar}. However, claiming that urban street networks universally are \enquote{planar} misrepresents them in several ways:

@@

As George Box once said, \enquote{All models are wrong but some are useful.} Even if they are not formally planar, can street networks be simplified to planar graphs and still be usefully well-modeled? Our results suggest that the answer depends on the study site and on the type of analysis. In limited circumstances, where the circulation network and built form exhibit few (or ideally zero) underpasses, overpasses, or grade-separation, then perhaps yes.

@@

We might repurpose this discrepancy between planar and nonplanar representations of a street network to reveal something useful about urban form. The $\phi$ and $\lambda$ scores of the major US city centers indicate the greater three-dimensionality of their transportation infrastructure compared to that of the European city centers. These US city centers feature more automobile-oriented roadways and are often bounded by grade-separated freeways. By contrast, these European city centers feature less three-dimensionality as their streets and paths tend to be at-grade. Examining how $\phi$ and $\lambda$ change over time in different cities could tell us something about the type and pace of urban development.

It is also interesting that most street networks are formally nonplanar while demonstrating $\phi$ and $\lambda$ scores that are only slightly nonplanar. Approximate planarity is common across cities because of costs, technology, and politics. Why? It is expensive to engineer a three-dimensional network with z-axis extents (nearly) as extensive as the x- and y-axes. Furthermore, a city fully consumed by overpasses and tunnels would present degraded livability and walkability, making it politically infeasible.

@@

%\clearpage
\bibliographystyle{apalike}
\bibliography{references}
